Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging

### 2022-10-26 analysis/paper/001-reply-main.Rmd
Minor textual changes to the second manuscript.

---
title: 'Supplementary Information to: "Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---

>>>>

---
title: 'Supplementary Information to: "Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
fontsize: 12pt
---

### 2022-10-26 analysis/paper/002-reply-main.Rmd
Minor textual changes to the second manuscript.

---
title: 'Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---

@@

2. The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased. An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e. holocellulose and Klason lignin) therefore is not possible.  
Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance. However, interpretation and generalization of these findings are less straightforward.

@@

In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a].  
The peat samples are more decomposed than the training data samples. Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a]. Since these features are not present in the training data, predictions of the original models are extrapolations. Our re-evaluation showed that this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >30 mass-% [@Teickner.2022a]. Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.  
Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of the quantitative concepts (i.e. holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat: as "carbohydrates" and "aromatics" (e.g. @Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014). We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.  
The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents. Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].   
An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry? For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?  
A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss. This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g. @Biester.2014). For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.

## Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}

>>>>

---
title: 'Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
fontsize: 12pt
---

@@

2. The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased. An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e. holocellulose and Klason lignin) is not possible.  
Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance. However, interpretation and generalization of these findings are less straightforward.

@@

In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a].  
The peat samples are more decomposed than the training data samples. Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a]. Since these features are not present in the training data, predictions of the original models are extrapolations. Our re-evaluation showed that this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >30 mass-% [@Teickner.2022a]. Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.  
Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of the quantitative concepts (i.e. holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat: as "carbohydrates" and "aromatics" (e.g. @Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014). We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.  
The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents. Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].   
An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry? For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?  
A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss. This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g. @Biester.2014). For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if as large environmental gradients are considered as in @Hodgkins.2018.

## Klason lignin contents and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}
